% This is LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science,
% version 2.4 for LaTeX2e as of 16. April 2010
%

\documentclass{llncs}
\usepackage{epstopdf}
%
\include{preamble}
\include{apis}
\include{utils}
\include{hyphenation}
\include{mytables}
%
\begin{document}
%&&&&&&&&&&&&&&&

\title{A Service Composition Approach for Data-Centric Workflows}
%
\titlerunning{A Service Composition Approach for Data-Centric Workflows}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{
Carlos-Manuel L\'opez-Enr\'iquez\inst{2} \and V\'ictor Cuevas-Vicentt\'in\inst{3} \and Genoveva Vargas-Solar\inst{1,2}
\and Christine Collet\inst{2} \and Jos\'e-Luis \\Zechinelli-Martini\inst{4}
}
%
\authorrunning{L\'opez-Enr\'iquez et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{Carlos-Manuel L\'opez-Enr\'iquez, V\'ictor Cuevas-Vicentt\'in, Genoveva Vargas-Solar, Christine Collet, Jos\'e-Luis \\Zechinelli-Martini}
%
\institute{
CNRS \and Grenoble Institute of Technology\\ BP. 72, 38402, Saint Martin d'H\`{e}res Cedex, France\\
\and
University of California at Davis\\ One Shields Avenue, Davis CA 95616, USA\\ 
\and
Exhacienda Sta. Catarina M\'{a}rtir s/n 72820 \\ San Andr\'{e}s Cholula, Puebla, Mexico \\
\email{carlos.manuel.lopez@gmail.com},\email{victorcuevasv@gmail.com},\email{genoveva.vargas@imag.fr},\\
\email{christine.collet@grenoble-inp.fr},\email{joseluis.zechinelli@udlap.mx} 
}


\maketitle              % typeset the title of the contribution

\begin{abstract}

This paper presents our approach for specifying, transforming, and enacting data-centric workflows in a service-oriented environment. First, we propose ASASEL (Abstract State Machines Execution Language) a language and system based on the ASM formalism for specifying workflows consisting of on-demand and streaming data services that produce data, which are then processed by simple and composite computation services as well as by operations on complex values. Second, we present our initial characterization of a planning-based approach for the transformation of workflows. In particular, we address the exploration of parallelism through the workflow structure. Together, our language and enactment engine along with our transformation approach provide the foundation for a highly flexible mechanism for managing data-centric workflows, and which can also be adapted to other domains such as business processes or eScience.

\keywords{workflows, services, answer set planning, logic programming}
\end{abstract}
%

\input{introduction.tex}

\input{dataCentricWorkflows.tex}

\input{complexValuesDataModel.tex}

\input{workflowTransformation.tex}

\input{experiments.tex}

\input{systemImplementation.tex}

\input{relatedWork.tex}

\input{conclusions.tex}


%}
%	In this paper we presented the constraints for generating query workflows that model the plans of hybrid queries.
%	The constraints represent the semantics of query workflow activities that is that implement data operators for evaluating a hybrid query.
%	We used the action language DLV-K for modeling the planing problem of query workflow generation in order to get a first (and naive) search space generation towards the optimization of hybrid queries.
%	The results show that time complexity for query workflow generation is exponential. 
%      
%	The constraints must be extended in order to consider the continuous data operators for completely modeling hybrid queries.
%	These constraints can be taken to model a more sophisticated query workflow generation.
%	This implies the abstraction of data operators types and their associated constraints in a general form. 
%
%	Theoretically, query workflow generation can be modeled mathematically by taking the constraints presented here and modeling an objective function.
%	Additionally, the equivalence and coherence of query workflows must be proved.


	
%
% ---- Bibliography ----
%
%\bibliographystyle{alpha}
\bibliographystyle{splncs03}
\bibliography{library/library}

\end{document}
